---
title: "CITEseq Cord Blood"
author: "Dario Righelli"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    BiocStyle::html_document:
      toc_float: true
vignette: >
    %\VignetteIndexEntry{CITEseq Cord Blood}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
Package: SingleCellMultiModal
bibliography: ../inst/REFERENCES.bib
editor_options:
  chunk_output_type: console
---

# Installation 

```{r,eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("SingleCellMultiModal")
```


# Load libraries

```{r, include=TRUE, results="hide", message=FALSE, warning=FALSE}

library(MultiAssayExperiment)
library(SingleCellMultiModal)

```


# CITE-seq dataset

CITE-seq data are a combination of two data types extracted at the same
time from the same cell.  First data type is scRNA-seq data, while the second
one consists of about a hundread of antibody-derived tags (ADT).
In particular this dataset is provided by @stoeckius2017simultaneous.

## Downloading datasets

The user can see the available dataset by using the default options

```{r}

CITEseq(DataType="cord_blood", modes="*", dry.run=TRUE, version="1.0.0")

```

Or simply by setting `dry.run = FALSE` it downloads the data and creates the
`MultiAssayExperiment` object.

In this example, we will use one of the two available datasets `scADT_Counts`:

```{r,message=FALSE}

mse <- CITEseq(
    DataType="cord_blood", modes="*", dry.run=FALSE, version="1.0.0"
)

mse
```

Example with actual data:

```{r}
experiments(mse)
```


## Exploring the data structure

Check row annotations:

```{r}
rownames(mse)
```

Take a peek at the `sampleMap`:

```{r}
sampleMap(mse)
```


## scRNA-seq data

The scRNA-seq data are accessible with the name `scRNAseq`, which returns a
*matrix* object.

```{r}
head(experiments(mse)$scRNAseq)[, 1:4]
```

## scADT data

The scADT data are accessible with the name `scADT`, which returns a
**matrix** object.

```{r}
head(experiments(mse)$scADT)[, 1:4]
```

# SingleCellExperiment object conversion

Because of already large use of some methodologies (such as
in the [SingleCellExperiment vignette][1] or [CiteFuse Vignette][2] where the
`SingleCellExperiment` object is used for CITE-seq data,
we provide a function for the conversion of our CITE-seq `MultiAssayExperiment`
object into a `SingleCellExperiment` object with scRNA-seq data as counts and
scADT data as `altExp`s.

[1]: https://www.bioconductor.org/packages/release/bioc/vignettes/SingleCellExperiment/inst/doc/intro.html#5_adding_alternative_feature_sets
[2]: http://www.bioconductor.org/packages/release/bioc/vignettes/CiteFuse/inst/doc/CiteFuse.html

```{r message=FALSE}
sce <- CITEseq(DataType="cord_blood", modes="*", dry.run=FALSE, version="1.0.0",
              DataClass="SingleCellExperiment")
sce
```

# Session Info

```{r, tidy=TRUE}
sessionInfo()
```

# References

