---
title: "scNMT Mouse Gastrulation"
date: "`r BiocStyle::doc_date()`"
vignette: |
  %\VignetteIndexEntry{scNMT Mouse Gastrulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
    BiocStyle::html_document:
      toc_float: true
Package: SingleCellMultiModal
bibliography: ../inst/REFERENCES.bib
---

# Installation

```{r,eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("SingleCellMultiModal")
```

## Load

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(SingleCellMultiModal)
library(MultiAssayExperiment)
```

# scNMT

The dataset was provided by Argelaguet and colleagues [@Argelaguet2019-et].

## Downloading datasets

The user can see the available dataset by using the default options

```{r}
scNMT("mouse_gastrulation", mode = "*", dry.run = TRUE)
```

Or by simply running:

```{r}
scNMT("mouse_gastrulation")
```

## Chromatin Accessibility

```{r}
nmt <- scNMT("mouse_gastrulation", mode = c("*_DHS", "*_cgi", "*_genebody"), dry.run = FALSE)
```

Check row annotations:

```{r}
rownames(nmt)
```

Take a peek at the `sampleMap`:

```{r}
sampleMap(nmt)
```

See the accessibilty levels (as proportions) for DNase Hypersensitive Sites:

```{r}
head(assay(nmt, "acc_DHS"))[, 1:4]
```

## DNA Methylation

See the methylation percentage / proportion:

```{r}
head(assay(nmt, "met_DHS"))[, 1:4]
```

For protocol information, see the references below.

# sessionInfo

```{r}
sessionInfo()
```

# References
