---
output: github_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# Overview

Thank you for your interest!

The `SingleCellMultiModal` package aims to provide datasets that have similar
pre-processing techniques from several different technologies. In particular,
our initial focus is on providing data from `scNMT` and `scM&T` technologies
but all multi-modal technologies are welcome.

# List of Multi-modal Datasets

Please see this spreadsheet for a list of available datasets:
[Google Drive Sheet](https://drive.google.com/drive)

# Contributing

In order to contribute, we require data in a specific format.
Aside from the usual required `metadata.csv` documentation in
the package, contributors are required to add a name to the
`DataType` column in the metadata table that indicates the name of the
contributed dataset.

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(SingleCellMultiModal)
```

```{r}
meta <- system.file("extdata", "metadata.csv",
    package = "SingleCellMultiModal", mustWork = TRUE)
head(read.csv(meta))
```

# vignettes

Contributors are required to demonstrate user-level functionality via
examples in a vignette for each contributed dataset.

# R

Ideally, the interface for the contributed dataset should be similar to that
of `scNMT` so that users have a sense of consistency in the usage of the
package. This means having one main function that returns a
`MultiAssayExperiment` object and having options that show the user what
datasets are available for a particular technology. Contributors should use
`roxygen2` for documenting datasets.

See the current example for implementation details:

```{r}
scNMT(dataType = "mouse_gastrulation", mode = "*", dry.run = TRUE)
```

# inst/*

## extdata/

The data should be organized in a data directory (`directory` argument in
`make_metadata`) that includes another folder with the name of the dataset
(`dataDir` in `make_metadata`) and documented in the `metadata.csv` under the
`DataType` name. The fields required by `ExperimentHub` are listed here:

* ResourceName
* Title
* Description
* BiocVersion
* Genome
* SourceType
* SourceUrl
* SourceVersion
* Species
* TaxonomyId
* Coordinate_1_based
* DataProvider
* Maintainer
* RDataPath
* RDataClass
* DispatchClass
* DataType

To make it easy for contributions, we've provided a mechanism for easy
documentation using a file from a `data.frame`.

Interested contributors should modify the file in the location here:

```{r}
docu <- system.file("extdata", "docuData", "singlecellmultimodalv2.csv",
    package = "SingleCellMultiModal", mustWork = TRUE)
docu
```

Contributors should add an additional row that documents the contributed
dataset. The example below shows the currently present datasets
(`dataType = "mouse_gastrulation"`):

```{r}
read.csv(docu, header = TRUE)
```

Contributors can create an entry to `singlecellmultimodal.csv` as a
`data.frame` and add the row. In this example, we create the data already
contained in the `.csv` file:

```{r}
scmeta <- data.frame(
    DataProvider =
        "Dept. of Bioinformatics, The Babraham Institute, United Kingdom",
    TaxonomyId = "10090",
    Species = "Mus musculus",
    SourceUrl = "https://cloudstor.aarnet.edu.au/plus/s/Xzf5vCgAEUVgbfQ",
    SourceType = "RDS",
    SourceVersion = c("1.0.0", "2.0.0"),
    DataType = "mouse_gastrulation",
    Maintainer = "Marcel Ramos <marcel.ramos@roswellpark.org>",
    stringsAsFactors = FALSE
)
scmeta
```

If the contributor can then append the metadata to the existing file.

### Saving the data

After creating the documentation `data.frame`, the contributor can save the
dataset as a `.csv` file using `write.csv` and append to the existing
documentation file.

```{r,eval=FALSE}
write.csv(
    scmeta,
    file = "inst/extdata/docuData/singlecellmultimodal.csv",
    row.names = FALSE
)
```

## Documenting diverse data

In the case that the contributed data is not uniform, meaning that there are
multiple file types in the data type folder, the `data.frame` will have to
account for _all_ contributed data files.
For example, if the contributed data has a number of different source types,
the contributor is required to create a `data.frame` with the number of rows
equal to the number of files to be uploaded.

In this example, we have two datasets from different source types and formats:

```{r}
data.frame(
    DataProvider = c("Institute of Population Genetics",
        "Genetic Science Center"),
    TaxonomyId = "9606",
    Species = "Homo sapiens",
    SourceUrl = c("https://science/org", "https://science/gov"),
    SourceType = c("RDS", "XML"),
    dataType = "human_genetics",
    stringsAsFactors = FALSE
)
```

## scripts/

Once the data is ready, the user can use the function in `make-metadata.R`
in the `scripts` folder.  A typical call to `make_metadata` will either add to
the metadata or replace it entirely. The easiest for current contributors is to
`append` rows to the metadata file.

```{r,eval=FALSE}
make_metadata(
    directory = "~/data/",
    dataDirs = "mouse_gastrulation", # also the name of the DataType
    ext_pattern = "\\.[Rr][Dd][Aa]$",
    doc_file = "../inst/extdata/docuData/singlecellmultimodalv3.csv",
    pkg_name = "SingleCellMultiModal",
    append = TRUE,
    dry.run = TRUE
)
```

The user may also wish to run `dry.run=TRUE` to see the output `data.frame`
to be added to the `metadata.csv` file.

```{r,eval=FALSE}
make_metadata(
    directory = "~/data/",
    dataDirs = "contributed_data", # also the name of the DataType
    ext_pattern = "\\.[Rr][Dd][Aa]$",
    doc_file = "../inst/extdata/docuData/mymeta.csv",
    pkg_name = "SingleCellMultiModal",
    append = TRUE,
    dry.run = TRUE
)
```

_Note_. The `make_metadata` function should be run from the base package
directory from a GitHub / git checkout (`git clone ...`).

The contributor should then create a Pull Request on [GitHub][].

[GitHub]: https://github.com/waldronlab/SingleCellMultiModal/pulls

If you are interested in contributing, I can help you go over the contribution
and submission. Please contact me either on the [Bioc-community Slack][]
(mramos148) or at marcel {dot} ramos [at] sph (dot) cuny (dot) edu.
If you need to sign up to the community Slack channel, follow this link:
https://bioc-community.herokuapp.com/

[Bioc-community Slack]: https://community-bioc.slack.com

## sessionInfo

<details>
  <summary> sessionInfo </summary>

```{r,echo=FALSE}
sessionInfo()
```

</details>
